Analysis of cloud enhancement events in a 30-year record of global solar irradiance at Thessaloniki, Greece

### 2024-09-02 article/article.Rmd
Commit 2024-09-02 09:09

author:
  - name:        Athanasios N. Natsis
    email:       natsisphysicist@gmail.com
    affiliation: LAP
    correspondingauthor: true
    # footnote: 1
  - name:        Alkiviadis Bais
    email:       abais@auth.gr
    affiliation: LAP
  - name:        Charikleia Meleti
    email:       meleti@auth.gr
    affiliation: LAP
    # footnote: 1
address:
  - code:         LAP
    organization: Laboratory of Atmospheric Physics, Physics Department, Aristotle University of Thessaloniki
    # addressline:  1 main street
    city:         Thessaloniki
    postcode:     54124
    country:      Greece
  # - code:         Another University
  #   organization: Department
  #   addressline:  A street 29
  #   postcode:     2054 NX
  #   city:         Manchester,
  #   state:        State
  #   country:      The Netherlands
# footnote:
#   - code: 1
#     text: "This is the first author footnote."
#   - code: 2
#     text: "Another author footnote."
abstract: |
  In this study we investigate the characteristics of global horizontal irradiance
  enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
  using data recorded every one minute. We identified the cloud enhancement (CE)
  events by creating an appropriate clear-sky irradiance reference with the use of a
  radiative transfer model and aerosol optical depth data from a collocated Cimel sun
  photometer and a Brewer spectrophotometer. We found a trend in CE events of
  \(45.6\pm 21.9\,\text{cases}/\text{year}\), and a trend in the CE events
  irradiation of \(116.9\pm 67.8\,\text{kJ}/\text{year}\). The peak of the CE events
  was observed during May and June. The analysis of the total duration of CE events
  showed that durations longer than 5 minutes are very rare, with exceptions lasting
  over an hour and up to 140 minutes. Finally, we have detected enhancements above
  the total solar irradiance at the top of the atmosphere of up to
  \(400\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the cases below
  \(200\,\text{W}/\text{m}^{2}\). The most active period of these extreme events is
  spring -- early summer with a secondary peak in autumn.

@@

<!-- Define CE and ECE tomove? -->
In this study, we define an event as CE when the measured GHI at ground level,
exceeds the expected value under clear-sky conditions. Similarly, we define as
extreme cloud enhancement events (ECE), the cases when GHI at ground level exceeds
the Total Solar Irradiance (TSI). Although the duration of these bursts can vary from
seconds to several minutes, here we are constrained by the temporal resolution of our
data to identify events with duration of at minimum one-minute.

@@

Following the application of the above discussed methodology to the entire dataset 
($`r round(ST_total$wattGLB.N / 10^6,1)`$ million of one-minute GHI measurements),
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 2)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(MC[Typ == "Max Glob", wattGLB],1)`\,\text{W}/\text{m}^2$ on
 `r strftime(MC[Typ == "Max Glob", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Glob", SZA],1)`^\circ$ corresponding to OI of
$`r round(MC[Typ == "Max Glob", GLB_diff],1)`\,\text{W}/\text{m}^2$ or
$`r round(100*MC[Typ == "Max Glob", GLB_ench],1)`\,\%$ above the threshold.
The strongest CE event of
$`r round(100*MC[Typ == "Max Enha", GLB_ench],1)`\,\%$ above the 
clear sky threshold was observed on 
`r strftime(MC[Typ == "Max Enha", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Enha", SZA],1)`^\circ$ with a GHI value of 
$`r round(MC[Typ == "Max Enha", wattGLB],1)`\,\text{W}/\text{m}^2$ and a OI of
$`r round(MC[Typ == "Max Enha", GLB_diff],1)`\,\text{W}/\text{m}^2$.
Both cases are ECE events with
$`r round(MC[Typ == "Max Glob", wattGLB - ETH],1)`\,\text{W}/\text{m}^2$ and
$`r round(MC[Typ == "Max Enha", wattGLB - ETH],1)`\,\text{W}/\text{m}^2$ above the
TSI at TSI for the same SZA, respectively.
In the following sections we are discussing the long-term trends and variability of
the CE events as well as of the corresponding OI and excess irradiation.

@@

The relation between the duration and the mean OI of the groups has also been
studied (Figure\nobreakspace{}\ref{fig:group-2d}).  Evidently, events of high excess
irradiation have small duration and vice versa. The vast majority of grouped events
are associated with small excess irradiation (e.g., $<5\,\text{kJ}/\text{m}^2$) and
small duration (e.g., $<5\,\text{min}$) while groups with strong excess irradiation
and long duration are very rare.  Similar results of this relation have been reported
by @Zhang2018 in a study using a far higher sampling rate ($100\,\text{Hz}$) than
ours.

@@

<!-- ```{r echo=FALSE, fig.cap="Relation of mean OI and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-4.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Relation of maximun OI and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-8.png") -->
<!-- ``` -->

@@

<!-- ```{r echo=FALSE, fig.cap="Distribution of the number of ECE cases"} -->
<!-- knitr::include_graphics("../images/climECEmonth-1.png") -->
<!-- ``` -->

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in rare cases the
OI can exceed the TSI even by more than $400\,\text{W}/\text{m}^2$, while in
$75\,\%$ of the cases the OI is below $200\,\text{W}/\text{m}^2$. The OI for the
most frequent ECEs ranges between $140$ and $180\,\text{W}/\text{m}^2$.  These
findings are in accordance with the results of @Vamvakas2020, the only difference
being that the OI values reported for are higher than those for Thessaloniki.

@@

<!-- ```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1994 -- 2023."} -->
<!-- knitr::include_graphics("../images/P-extreme-distribution-1.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Distribution of ECE Irradiance above 'TSI'"} -->
<!-- knitr::include_graphics("../images/extremedistributions-4.png") -->
<!-- ``` -->

<!--  SZA -->

<!-- ### Yearly plots relative to all GHI data. -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-1.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-2.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-3.png") -->
<!-- ``` -->

>>>>

author:
  - name:        Athanasios N. Natsis
    email:       natsisphysicist@gmail.com
    affiliation: LAP
    correspondingauthor: true
    # footnote: 1
  - name:        Alkiviadis Bais
    email:       abais@auth.gr
    affiliation: LAP
  - name:        Charikleia Meleti
    email:       meleti@auth.gr
    affiliation: LAP
    # footnote: 1
address:
  - code:         LAP
    organization: Laboratory of Atmospheric Physics, Physics Department, Aristotle University of Thessaloniki
    # addressline:  1 main street
    city:         Thessaloniki
    postcode:     54124
    country:      Greece
  # - code:         Another University
  #   organization: Department
  #   addressline:  A street 29
  #   postcode:     2054 NX
  #   city:         Manchester,
  #   state:        State
  #   country:      The Netherlands
# footnote:
#   - code: 1
#     text: "This is the first author footnote."
#   - code: 2
#     text: "Another author footnote."
# abstract: |
#   In this study we investigate the characteristics of global horizontal irradiance
#   enhancement events induced by clouds over Thessaloniki for the period 1994 -- 2023
#   using data recorded every one minute. We identified the cloud enhancement (CE)
#   events by creating an appropriate clear-sky irradiance reference with the use of a
#   radiative transfer model and aerosol optical depth data from a collocated Cimel sun
#   photometer and a Brewer spectrophotometer. We found a trend in CE events of
#   \(45.6\pm 21.9\,\text{cases}/\text{year}\), and a trend in the CE events
#   irradiation of \(116.9\pm 67.8\,\text{kJ}/\text{year}\). The peak of the CE events
#   was observed during May and June. The analysis of the total duration of CE events
#   showed that durations longer than 5 minutes are very rare, with exceptions lasting
#   over an hour and up to 140 minutes. Finally, we have detected enhancements above
#   the total solar irradiance at the top of the atmosphere of up to
#   \(400\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the cases below
#   \(200\,\text{W}/\text{m}^{2}\). The most active period of these extreme events is
#   spring -- early summer with a secondary peak in autumn.

@@

<!-- Define CE and ECE tomove? -->
In this study, we define an event as CE when the measured GHI at ground level,
exceeds the expected value under clear-sky conditions. Similarly, we define as
extreme cloud enhancement events (ECE), the cases when GHI at ground level exceeds
the Total Solar Irradiance (TSI) at the top of the atmosphere (TOA) for the same SZA.
Although the duration of these bursts can vary from seconds to several minutes, here
we are constrained by the temporal resolution of our data to identify events with
duration of at minimum one-minute.

@@

Following the application of the above discussed methodology to the entire dataset 
($`r round(ST_total$wattGLB.N / 10^6,1)`$ million of one-minute GHI measurements),
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 2)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(MC[Typ == "Max Glob", wattGLB],1)`\,\text{W}/\text{m}^2$ on
 `r strftime(MC[Typ == "Max Glob", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Glob", SZA],1)`^\circ$ corresponding to OI of
$`r round(MC[Typ == "Max Glob", GLB_diff],1)`\,\text{W}/\text{m}^2$ or
$`r round(100*MC[Typ == "Max Glob", GLB_ench],1)`\,\%$ above the threshold.
The strongest CE event of
$`r round(100*MC[Typ == "Max Enha", GLB_ench],1)`\,\%$ above the 
clear sky threshold was observed on 
`r strftime(MC[Typ == "Max Enha", Date], "%d\\ %B %Y")` at a SZA of
$`r round(MC[Typ == "Max Enha", SZA],1)`^\circ$ with a GHI value of 
$`r round(MC[Typ == "Max Enha", wattGLB],1)`\,\text{W}/\text{m}^2$ and a OI of
$`r round(MC[Typ == "Max Enha", GLB_diff],1)`\,\text{W}/\text{m}^2$.
Both cases are ECE events with
$`r round(MC[Typ == "Max Glob", wattGLB - ETH],1)`\,\text{W}/\text{m}^2$ and
$`r round(MC[Typ == "Max Enha", wattGLB - ETH],1)`\,\text{W}/\text{m}^2$ above the
TSI at TOA for the same SZA, respectively.
In the following sections we are discussing the long-term trends and variability of
the CE events as well as of the corresponding OI and excess irradiation.

@@

The relation between the duration and the mean OI of the groups has also been studied
(Figure\nobreakspace{}\ref{fig:group-2d}).  Evidently, events of high excess
irradiation have small duration and vice versa. The vast majority of grouped events
are associated with small excess irradiation ($<5\,\text{kJ}/\text{m}^2$) and small
duration ($<5\,\text{min}$). For example the extra energy flux of
$5\,\text{kJ}/\text{m}^2$ for $5\,\text{min}$ is equivalent to OI of
$\sim17\,\text{W}/\text{m}^2$ for the same duration. While groups with strong excess
irradiation and long duration are very rare.  Similar results of this relation have
been reported by @Zhang2018 in a study using a far higher sampling rate
($100\,\text{Hz}$) than
ours.

@@

@@

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in rare cases the
OI can exceed the TSI at TOA for the same SZA even by more than
$400\,\text{W}/\text{m}^2$, while in $75\,\%$ of the cases the OI is below
$200\,\text{W}/\text{m}^2$. The OI for the most frequent ECEs ranges between $140$
and $180\,\text{W}/\text{m}^2$.  These findings are in accordance with the results of
@Vamvakas2020, the only difference being that the OI values reported for are higher
than those for Thessaloniki.

@@

### 2024-09-02 article/article.tex
Commit 2024-09-02 09:09

  \begin{abstract}
  In this study we investigate the characteristics of global horizontal
  irradiance enhancement events induced by clouds over Thessaloniki for
  the period 1994 -- 2023 using data recorded every one minute. We
  identified the cloud enhancement (CE) events by creating an
  appropriate clear-sky irradiance reference with the use of a radiative
  transfer model and aerosol optical depth data from a collocated Cimel
  sun photometer and a Brewer spectrophotometer. We found a trend in CE
  events of \(45.6\pm 21.9\,\text{cases}/\text{year}\), and a trend in
  the CE events irradiation of \(116.9\pm 67.8\,\text{kJ}/\text{year}\).
  The peak of the CE events was observed during May and June. The
  analysis of the total duration of CE events showed that durations
  longer than 5 minutes are very rare, with exceptions lasting over an
  hour and up to 140 minutes. Finally, we have detected enhancements
  above the total solar irradiance at the top of the atmosphere of up to
  \(400\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the cases below
  \(200\,\text{W}/\text{m}^{2}\). The most active period of these
  extreme events is spring -- early summer with a secondary peak in
  autumn.
  \end{abstract}
    \begin{keyword}
    cloud enhancement \sep total solar radiation \sep global horizontal
irradiance \sep 
    over irradiance
  \end{keyword}

@@

In this study, we define an event as CE when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions. Similarly,
we define as extreme cloud enhancement events (ECE), the cases when GHI
at ground level exceeds the Total Solar Irradiance (TSI). Although the
duration of these bursts can vary from seconds to several minutes, here
we are constrained by the temporal resolution of our data to identify
events with duration of at minimum one-minute.

@@

Following the application of the above discussed methodology to the
entire dataset (\(6\) million of one-minute GHI measurements),
\(1.78\,\%\) were identified as CE events and \(0.04\,\%\) as ECE
events. The highest recorded GHI due to CE was
\(1416.6\,\text{W}/\text{m}^2\) on 24~May 2007 at a SZA of
\(19.9^\circ\) corresponding to OI of \(345.9\,\text{W}/\text{m}^2\) or
\(32.3\,\%\) above the threshold. The strongest CE event of \(49.7\,\%\)
above the clear sky threshold was observed on 28~October 2016 at a SZA
of \(59.2^\circ\) with a GHI value of \(861.8\,\text{W}/\text{m}^2\) and
a OI of \(286.1\,\text{W}/\text{m}^2\). Both cases are ECE events with
\(136.8\,\text{W}/\text{m}^2\) and \(164.5\,\text{W}/\text{m}^2\) above
the TSI at TSI for the same SZA, respectively. In the following sections
we are discussing the long-term trends and variability of the CE events
as well as of the corresponding OI and excess irradiation.

@@

The relation between the duration and the mean OI of the groups has also
been studied (Figure\nobreakspace{}\ref{fig:group-2d}). Evidently,
events of high excess irradiation have small duration and vice versa.
The vast majority of grouped events are associated with small excess
irradiation (e.g., \(<5\,\text{kJ}/\text{m}^2\)) and small duration
(e.g., \(<5\,\text{min}\)) while groups with strong excess irradiation
and long duration are very rare. Similar results of this relation have
been reported by \citet{Zhang2018} in a study using a far higher
sampling rate (\(100\,\text{Hz}\)) than ours.

@@

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in
rare cases the OI can exceed the TSI even by more than
\(400\,\text{W}/\text{m}^2\), while in \(75\,\%\) of the cases the OI is
below \(200\,\text{W}/\text{m}^2\). The OI for the most frequent ECEs
ranges between \(140\) and \(180\,\text{W}/\text{m}^2\). These findings
are in accordance with the results of \citet{Vamvakas2020}, the only
difference being that the OI values reported for are higher than those
for Thessaloniki.

>>>>

  \begin{abstract}

  \end{abstract}
    \begin{keyword}
    cloud enhancement \sep total solar radiation \sep global horizontal
irradiance \sep 
    over irradiance
  \end{keyword}

@@

In this study, we define an event as CE when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions. Similarly,
we define as extreme cloud enhancement events (ECE), the cases when GHI
at ground level exceeds the Total Solar Irradiance (TSI) at the top of
the atmosphere (TOA) for the same SZA. Although the duration of these
bursts can vary from seconds to several minutes, here we are constrained
by the temporal resolution of our data to identify events with duration
of at minimum one-minute.

@@

Following the application of the above discussed methodology to the
entire dataset (\(6\) million of one-minute GHI measurements),
\(1.78\,\%\) were identified as CE events and \(0.04\,\%\) as ECE
events. The highest recorded GHI due to CE was
\(1416.6\,\text{W}/\text{m}^2\) on 24~May 2007 at a SZA of
\(19.9^\circ\) corresponding to OI of \(345.9\,\text{W}/\text{m}^2\) or
\(32.3\,\%\) above the threshold. The strongest CE event of \(49.7\,\%\)
above the clear sky threshold was observed on 28~October 2016 at a SZA
of \(59.2^\circ\) with a GHI value of \(861.8\,\text{W}/\text{m}^2\) and
a OI of \(286.1\,\text{W}/\text{m}^2\). Both cases are ECE events with
\(136.8\,\text{W}/\text{m}^2\) and \(164.5\,\text{W}/\text{m}^2\) above
the TSI at TOA for the same SZA, respectively. In the following sections
we are discussing the long-term trends and variability of the CE events
as well as of the corresponding OI and excess irradiation.

@@

The relation between the duration and the mean OI of the groups has also
been studied (Figure\nobreakspace{}\ref{fig:group-2d}). Evidently,
events of high excess irradiation have small duration and vice versa.
The vast majority of grouped events are associated with small excess
irradiation (\(<5\,\text{kJ}/\text{m}^2\)) and small duration
(\(<5\,\text{min}\)). For example the extra energy flux of
\(5\,\text{kJ}/\text{m}^2\) for \(5\,\text{min}\) is equivalent to OI of
\(\sim17\,\text{W}/\text{m}^2\) for the same duration. While groups with
strong excess irradiation and long duration are very rare. Similar
results of this relation have been reported by \citet{Zhang2018} in a
study using a far higher sampling rate (\(100\,\text{Hz}\)) than ours.

@@

The distribution of the ECE events
(Figure\nobreakspace{}\ref{fig:P-extreme-distribution}) shows that in
rare cases the OI can exceed the TSI at TOA for the same SZA even by
more than \(400\,\text{W}/\text{m}^2\), while in \(75\,\%\) of the cases
the OI is below \(200\,\text{W}/\text{m}^2\). The OI for the most
frequent ECEs ranges between \(140\) and \(180\,\text{W}/\text{m}^2\).
These findings are in accordance with the results of
\citet{Vamvakas2020}, the only difference being that the OI values
reported for are higher than those for Thessaloniki.
